### 2017-07-14 sheets/talk-11/sheet-11.tex

\begin{hint}
  Read Martin Hofmann's Syntax and Semantics of Dependent Types and section $1$
  in Simon Huber's A Model of Type Theory in Cubical Sets.
\end{hint}

\begin{exercise}
  Read Hofmann's section on type for natural numbers (section 2.1.3).

@@

  Define a length function of type $List(A) \to \mathbb{N}$ and
  define a type $Vec_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
  using lists, the identity type and the $\Sigma$-type.
\end{exercise}

@@

\begin{exercise}
  Recall the notation for the category of elements of a presheaf $\Gamma$:
  $\int_\cat{C}\Gamma$ Check that substitution of $A\from
  (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from \Delta \to \Gamma$
  corresponds to precomposing with $\int_\cat{C} \sigma : \int_\cat{C} \Delta
  \to \int_\cat{C} \Gamma$ induced by $\sigma$.

\begin{exercise}
  Define dependent sum using presheaf model notation

  \emph{Hint}: Read section $1.2$ in Huber's thesis to recall presheaf model
  construction. Try to come up with your definition before checking the answer
  given in the section.
\end{exercise}
\begin{answer}
  Section $1.2.2$.
\end{answer}

>>>>

\begin{hint}
  Read sections $2.1.3, 2.1.5, 2.1.6$ in Martin Hofmann's Syntax and Semantics
  of Dependent Types and section $1$ in Simon Huber's A Model of Type Theory in
  Cubical Sets.
\end{hint}

\begin{exercise}
  Describe the elements of $Ty_F(\Gamma)$ and $Ter_F(\Gamma, A)$.
\end{exercise}
\begin{answer}
  $Ty_F(\Gamma)$ is the first projection in $F(\Gamma)$, that is a set with some
  objects (possibly empty).
  \[
    Ty_F(\Gamma) := \{A_\Gamma, B_\Gamma, \dots\}
  \]
  $Ter_F(\Gamma, A)$ is then an element of the second projection of $F(\Gamma)$,
  that is a member of the family of sets indexed over $Ty_F(\Gamma)$, which has
  A as its index and also may be empty.
  \[
    Ter_F(\Gamma, A) := \{a_A, b_A, \dots\}_A
  \]
\end{answer}

\begin{exercise}
  Check that the following square is a pullback in Categories with Families.
  \[
    \begin{tikzcd}
      \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
      \Delta \arrow{r}{\sigma} & \Gamma
    \end{tikzcd}
  \]
  where $\Delta.A \vdash q : A p_\Delta$
\end{exercise}

\begin{exercise}
  Recall the notation for the category of elements of a presheaf $\Gamma$
  (section $1.2$ p. $16$ in Huber's work): $\int_\cat{C}\Gamma$ Check that
  substitution of $A\from (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from
  \Delta \to \Gamma$ corresponds to precomposing with $\int_\cat{C} \sigma :
  \int_\cat{C} \Delta \to \int_\cat{C} \Gamma$ induced by $\sigma$.

\begin{exercise}
  From the slides we know that If $B$ in $\Pi x: A. B$ does not depend on $A$,
  then $\Pi x: A. B$ corresponds to the ordinary non-dependent function type:
  \begin{equation*}
    A \to B \vcentcolon= \Pi x: A. B
  \end{equation*}

  Can you give a term of the type $A \to A$ using the presheaf model?

  \emph{Hint}: Read section $1.2.1$ in Huber's work for the general case.

\end{exercise}

\begin{exercise}
  Suppose we have two independent of each other types $\Gamma \vdash A \text{
  type}$ and $\Gamma \vdash B \text{ type}$.
  Then we can define a product type in our presheaf model in the following way:
  \[
    (A \times B)\rho := \{(a,b) | a \in A\rho \land b \in B\rho\}
  \]

  Now suppose that we have dependent types $\Gamma \vdash C \text{ type}$ and
  $\Gamma, C \vdash D \text{ type}$.

  How would you define in the presheaf model a product type that involves a
  dependent type (i.e. the dependent sum type)?

  \[
    (\Sigma AB)\rho := ???
  \]

  \emph{Hint}: The definition is \emph{almost} identical to the one given
  above. But what do you need to change? To recall the presheaf model, read
  section $1.2$ in Huber's work.
\end{exercise}
\begin{answer}
  Section $1.2.2$.
\end{answer}

\begin{exercise}
  Read Hofmann's section on type for the natural numbers (section 2.1.3).

@@

  Define a length function of type $List(A) \to \mathbb{N}$ and
  define a type $\mathit{Vec_A}(n)$ of lists of length $n$ for each $n : \mathbb{N}$
  using lists, the identity type and the $\Sigma$-type.
\end{exercise}

@@
